The Galactic Chemical Evolution of Carbon: Implications for Stellar Nucleosynthesis

### 2023-07-21 carbon_paper/tex/main.tex

% Abstract of the paper
\begin{abstract}
% context
C is an important element across astronomy. However, its origin remains poorly understood. 
% 
As stellar yields are highly uncertain, we aim to constrain the stellar yields of C through multi-zone Galactic chemical evolution models. 
% results
We use \apogee\ subgiants to constrain the model. 
% 
We find that \caafe\ is an emirical estimate of the delayed C sources, enabling us to estimate that \agb\ stars and \cc\ produce about 20\% and 80\% of C, respectively.  
The \caah\ trend instead represents the equilibrium abundances of C and Mg. 
We use the \caah\ trend to estimate the CCSNe C/Mg yield, determining that  $\Ycc/\Yoc = EQUATION$ when including \agb\ C. 
% misc
Our models are relatively independent of uniform scaling of yields and outflows, and alternate star formation histories. 
However, the stars which contribute to \agb\ C production and the \ia\ delay time distribution of Fe contribute uncertainties to our conclusions. 
% gas phase
While reliable gas-phase and low-metallicity measurments of C are challenging, we find that our model and a singlezone model with our recommended yields replicate the broad trends of \caah across different environments and metallicities.

@@

Carbon is a distinctive and well-studied element in astronomy. 
Formed in the cores of stars during He fusion, C is the lightest directly synthesized element after He, one of the only light elements formed in low-mass stars, and one the most abundant metals\footnotemark{}. % \citep[e.g.][]{jennifer19, KL14}
C structurally changes the environments it pollutes, modifying stellar evolution, facilitating the formation of stars and planets, and forming the basis of earthly life.
% C/N as age indicator?
As such, understanding the origin of C has wide ranging implications. 
While we know both lower-intermediate-mass and high-mass stars produce C, the relative importance of each process is still unknown.

\footnotetext{Astronomers are not chemists. By metallicity, we mean the (mass) fraction of any element which is not H or He, denoted by $Z$. For the sun, $Z_\odot=0.014$. }

Galactic chemical evolution is a powerful tool in understanding the origin of elements. 
Each enrichment process has distinct chemical signatures and timescales, enabling us to reconstruct chemical histories.
Many previous works have used GCE in attemt to understand C abundances, whether through using theoretical stellar models \citep{DTS78, prantzos+18, chiappini+03} 
or understanding  observation \citep{tinsley79, HEK00, BF06, rybizki+17, berg+19, KKL20}.
(See also review in \citealt{romano22}.)
%
Every study agrees that C is produced by a combination of high-mass and \lims{}, different studies disagree on which process is dominant. 
For example CITES conclude \hms{} dominate wherease CITES conclude \lims{} contribute the majority of C.
C is also generally understood to have strong metalliicity dependent CCSNe enrichment.

One of the primary uncertainties of GCE models are nucleosynthetic yields. Yield predictions -- the amounts of each chemical element stars produce --
are shaped by poorly understood processes, including mass loss, nuclear reaction rates, rotational mixing, convection, and explodability \citep{romano+10,KL14,ventura+13, LC18, emily+21}.
To better understand where C comes from and how it evolves, our aim is to use a combination of \apogee\ observations and multi-zone models to develop estimates of the required C yields to reproduce galactic trends.
\cite{james+23} examined similar \gce{} models of N (which is closely related to C), finding that trends in N and O\footnotemark{} are explained by the metallicity dependence of N/O yields. \citet{james+23} determine that \agb\ N abundances roughly dependence on metallicity (i.e. $y_{\rm N}/y_{\rm O} \propto Z$). 
Here, we extend their models to C, deriving similar constraints on C/Mg yields. We assess which yield prescriptions reproduce Galactic abundance trends while investigating the impact of \gce{} model assumptions, such as the star formation history (\sfh{}) and outflow mass loading.

\footnotetext{e.g. \citet{HEK00,PVT10,berg+12, berg+20, skillman+20, izotov+12, james2+15, dopita+16}.} 
We know, from observations of MW stars and extragalactic gas measurments, that C/O traces a banana shape in O/H (see \ref{fig:gas_phase}). 
% \cite[e.g.][]{fabbian+09, nissen+14, lambert81, laird85, lambert86}. 
At very low metallicities, C/O decreases with increasing metallicity, from observations of halo stars and damped Lyman-alpha systems\footnotemark{} \citep{nissen+14, cooke+17, fabbian+09}. 
\footnotetext{DLAs; high redshift clouds of gas observed as absorption lines in yet higher redshift quasars}
%
However, the trend reverses around $[{\rm O/H}]\approx-1$\footnotemark{}. 
C/O instead increases with increasing metallicity in observations of HII regions and Milky Way stars (\citealt{berg+19}; see discussion in section \ref{sec:gas}).

@@

However, observations of C are not without challenges. 
As stars ascend the red giant branch (RGB) at the end of H core burning, C abundances are changed by \textit{first dredge up}\footnotemark{}. This has the benefit of allowing C/N to be used as an age indicator in RGB stars \citep{MG15, martig16, hasselquist19, vincenzo+21}, but evolved stars are no longer as useful for determining birth abundances. Works such as \citet{vincenzo+21} attempt to correct for these changes but this requires MESA stellar modeling. 
In optical spectroscopy, C lines are relatively faint, and surveys such as GALAH struggle against low detection rates, potentially biasing sample measurments. While \apogee\ includes more reliable C$_2$ lines, there is an additional risk of systematics because .....
In the gas phase, HII regions are our best window into C abundances in our galaxy and extragalactic regions including dwarf. 
Unfortunantly, C lacks strong optical emission lines. Approaches use either recombination lines or ... in the ultraviolet, but these methods disagree, making gas-phase abundances a limited metric for GCE models of C.

    As our primary observational constraint, we use a sample of subgiant stars from the \apogee{} \citep{apogee17}. selected by the criteria in Roberts et al. (2023, in prep; see Appendix \ref{sec:jack}).
According to stellar evolution theory and observations \citep{gilroy89, korn+07, lind+08, souto+18, souto19}, these stars have not yet experienced first dredge up (FDU) but have well-mixed envelopes. So, subgiant surface abundances should represent their birth composition. 
Fig.~\ref{fig:subgiants} shows the subgiant sample plotted in \caah\ and \caafe{}. [C/Mg] increases with metallicity, and [C/Mg] decreases with [Mg/Fe] at fixed [Mg/H]. 
Using these trends, we will develop a model of the enrichment sources and evolution of C.

\citet{james+23} compare their model against the \cite{vincenzo+21} sample of \apogee{} \citep{apogee17} RGB stars. \citet{vincenzo+21} use \textsc{mesa} stellar evolution models \citep{mesa} to correct for the effects of first dredge up. 
We instead use the \citetjack~sample of \apogee{} subgiants. Subgiant stars have not undergone first dredge up but have well-mixed atmospheres, so their atmospheric C and N abundances are still reflective of their birth abundances and do not need mixing corrections. We only compare the models to the low-$\alpha$ sequence unless otherwise specified.  The selection criteria and differences between the samples are described in more detail in Appendix \ref{sec:jack}.

\footnotetext{First dredge up (FDU) is when material from CNO processed core is mixed with the (convective) stellar envelop during the ascent onto the RGB. As a result, C is decreased and N is increased depending on the strength of this process and the envelop mass.}

@@

Yields -- the amounts of each chemical element stars synthesise -- are central to studies of galactic chemical evolution (GCE). 
In this section, we discuss the yield choices of our model and literature models for C yields. 
We focus on three nucleosynthetic pathways: asymptotic giant branch (\agb{}) stars, core collapse supernovae (\cc{}), and type Ia supernovae (\ia{}).
C is produced in both \agb{} and \cc{} stars.
We also use Mg and Fe as tracers of \cc{} and \ia{} enrichment respectively. 
O and Mg are produced almost entirely from \cc\ with metallicity-independent yields. In contrast Fe is produced in similar amounts by \cc\ and \ia.

@@

To quantify yields, we define the stellar yield to be the fraction of a stars initial mass which is newly synthesized and released as a given element. 
For an element $X$ and star with mass $M$, the net-fractional stellar yield $\y$ is 
\begin{equation}
\y_{X} =  \Delta Z_X \frac{M_{\rm ejected}}{M_{\rm birth}}
\end{equation}
where $M_{\rm ejected}$ and $M_{\rm birth}$  are the total ejected mass and the birth mass of the star, and $\Delta Z_X$ is the change in $Z_X$ from the birth material to the ejected material of the star.%
\footnote{$Z_{X}$ represents the mass fraction of element $X$.}
%
For example, a 1\,\Mo\ star with $\y_{\C} = 10^{-3}$ will add $10^{-3}\,\Mo$ of new C to the interstellar medium. 
Also, note that yields may be negative if the material returned to the interstellar medium has a lower abundance $Z_X$ than the material the star was formed from.
Although per-star yields are necessary to compute \agb{} star enrichment rates in \gce{}  models, \imf-averaged yields are useful in interpreting their predictions. 
Given the fractional yield $\y(M, Z)$ as a function of initial stellar mass $M$ and metallicity $Z$, the \imf-averaged yield is given by 
\begin{equation} \label{eq:imf-yield}
    y_{\rm X}(Z,t) = 
    \frac{
    \int_{M_{\rm min}(t)}^{M_{\rm max}} 
    \y_{\rm X}(M, Z)
    \frac{dN}{dM}\ M\ dM
}
{
    \int \frac{dN}{dM}\ M\ dM
}
\end{equation}
where ${dN}/{dM}$ is the \imf, $M_{\rm max}=100\,\Mo$ is the maximum stellar mass, and $M_{\rm min}(t)$ is the mass of stars with lifetime $t$.\footnotemark{}
We use $t=10\,$Gyr for total yields when $t$ is not used.
To calculate the \imf-averaged net yields, we use the Versatile Integrator for Chemical Evolution code (\VICE).\footnotemark{}

We adapt the yield choices of elements besides C from \citet{james+21, james+23}.
Table \ref{tab:fiducial_mod} contains our fiducial yields. 
Following \citet{james+21, james+23}, we also take the \ia{} delay time distribution to be a $t^{-1.1}$ power-law, as suggested by the observations of \citet{maoz+12}.

\addtocounter{footnote}{-2}
    \stepcounter{footnote}\footnotetext{In our model, the mass-lifetime relation is
$\log \tau_M = 1.02 - 3.57\log M + 0.90 \left(\log M\right)^2$,
where $\tau_M$ is in Gyr, from \citealt{larson74}}

@@

An \agb\ star is a low-mass star ($\lesssim 8 M_{\sun}$) during its final phase of evolution.  
In an \agb\ star, two competing processes determine the outcome of C production: \textit{third dredge up} and \textit{hot bottom burning}.  
Third dredge up accompanies thermal pulses in \agb\ stars, where material from the CO core is mixed with the envelope, increasing surface C abundances \citep{KL14}. The C yields of the star are increased as this C-enhanced envelope is released to the interstellar medium. 
Hot bottom burning is the activation of the CNO cycle\footnotemark{}
at the bottom of the convective envelope when $T\gtrsim 50\,{\rm MK}$. Because the $^{14}$N proton capture is the slowest component of the CNO cycle, the CNO cycle converts nearly all \C[12] into $^{14}$N \citep{solar-fusion}.

@@

Hot bottom burning and third dredge-up result in mass-dependent C yields. 
Stars less than \about{2}\,\Mo do not experience third dredge-up. As a result, these stars C abundances are only affected by first dredge-up, resulting in little change to C yields or slight destruction of C.
Between \about{2} and \about{5} \Mo, third dredge up becomes important, enriching the outer layers with C, making these stars the most important producers of C. 
In \agb\ stars more massive than \about{5}\,\Mo, both hot bottom burning and third dredge up occur; however, third dredge up is much more efficient, resulting in significant \C[12] destruction.

@@

Fig.~\ref{fig:y_agb} compares the stellar \agb\ C yield for these four models.
Most models agree on the qualitative shape of the net fractional \agb\ C yield.
Stellar yields peak between masses of about 2--4 \Mo and decline as stars become more or less massive. As metallicity increases, the total \agb\ C yield decreases. The mass of peak C yields also increases slightly with metallicity. More metal-rich stars lose mass more quickly due to higher opacities, so these stars behave more like low-mass AGB stars. MORE PHYSICS

Fig.~\ref{fig:agb-ssp}, on the left, shows the total production of C by \agb\ stars in a \ssp{} at an age $t$, i.e. $\y_{\rm C}(Z_\odot, t)$. 
As the mass range $2\,\Mo\lesssim M \lesssim 4\,\Mo$ is most important for C production, about half of C production occurs before \about{1}\,Gyr, similar to \ia\ Fe. 
\kx{} and \kxvi{} weight C production more heavily towards high-mass \agb\ stars resulting in a faster enrichment delay time, whereas the \cxi\ and \vxiii\ models predict a slightly longer timescale of \about{1}\,Gyr. In any case, little to no C is produced more than 2\,Gyr after a star formation event. In contrast, Fe production continues steadily for 10\,Gyr.

The right panel of Fig.~\ref{fig:agb-ssp} shows \imf-averaged C yields for each \agb\ model as a function of metallicity.
\vxiii{} differs in that it shows a non-monotonic metallicity dependence. However, this effect is only for models with $\log Z/Z_\odot \lesssim -1$.
Otherwise, models differ only in their yield normalization and metallicity dependence. 
All yield models span a range of \about{2} for a given metallicity.
For example, the three models \cxi, \kx{}, and \kxvi{} predict $y_\text{C}^\text{agb}$ to be between 0.006 and 0.008 at solar metallicity, but \cxi\ has a much shallower metallicity dependence than the \kx{} and \kxvi{} models. \vxiii{} instead predicts a yield \about{0.004}.

@@

Massive stars form $^{12}$C in their cores through the triple--$\alpha$ process. However, only C ejected through supernovae and stellar winds contributes to the yield. 
While there are many stellar models providing predictions of \cc{} yields, the results of these models are highly uncertain due to the complexity of stellar modeling.

@@

Both rotation and explodability introduce substantial variations. The \cite{LC18} models include rotation, showing that variations in the rotational velocity of the star can dramatically increase the magnitude and metallicity dependence of $\Ycc$. Rotation induces more mixing allowing the CO core to grow larger and contributes to wind losses. As we will later show, \cc\ C production needs to be strongly metallicity-dependent at $Z/Z_\odot \approx 1$, which is consistent with the \cite{LC18} rapidly rotating models.
Assumptions about the explodability landscape affect C and Mg production. Increasing the fraction of stars that explode increases $\Ycc$, as stars that directly collapse do not contribute to explosive yields \citep{emily+21}. However, C is relatively unaffected by the black-hole landscape, as very massive stars contribute C through enriched winds. Since Mg is formed deeper in the core of \hms, the Mg yield drops much more steeply, so models where few stars explode (S16/W18) have higher [C/Mg].

O and Mg are both $\alpha$-elements, light elements produced in \cc{} through He-nuclei fusion. 
As we focus on constraining relative yields, we neglect O and Mg yield variations in the main text (excluding the uniform scaling of yields and massloading in Section~\ref{sec:outflows}). There is substantial variation in predicted Mg yields. 
Most models predict relatively flat trends metallicity (even with rotation as in \citealt{LC18}). 
However, the variation is significant and our adopted $\Yoc$ yield is much higher than most models, but O and Mg yields of \cc\ models do not fully match observations. \cc\ models underpredict $[{\rm Mg}/{\rm O}]$, and the reason why is unknown (see e.g. \citealt{emily+21}). Here, we assume ${\rm [O/Mg]} = 0$, which is not compatible with \cc\ models but is consistent with \apogee{} observations \citep{weinberg+19, weinberg+22}.
Because O and Mg are not perfectly identical in \apogee{}, the specific choice of $\alpha$-element may be a possible systematic.

@@

Here, we assume a simple \textit{one-zone} chemical evolution model \cite[e.g.][]{tinsley80, pagel09, matteucci21}.  Newly produced metals are homogeneously and instantaneously mixed, so this model evolves as a single, spatially-uniform chemical zone.
We define $M_{X}$ to be the mass of element $X$ in the gas-phase, $\dot{M}_\star$ to be the star formation rate (in $\Mo\,{\rm yr}^{-1}$), and $\eta$ to be the mass loading factor $\eta\equiv\dot{M}_{\rm outflow}/\dot{M}_\star$ (representing the strength of outflows). A \ssp\ returns a fraction $r$ of their birth mass to the interstellar medium, due to ejected stellar envelopes.%
\footnote{$r\approx0.4$ for a \citealt{kroupa01} \imf.}
Here, we will use $\alpha$ as a representative $\alpha$-element, such as O or Mg. 
Given the \imf-averaged yield $y_\alpha$, the rate of change in the gas-phase mass of $\alpha$ is a simple sum of sources and sinks,
\begin{equation} \label{eq:mdot_1}
    \dot{M}_\alpha =  y_\alpha\,\dot{M}_\star - \dot{M}_{\alpha,\,\text{remnants}} - \dot{M}_{\alpha,\,\text{outflows}}
\end{equation}
where $y_\alpha\,\dot{M}_{\star}$ describes \cc\ enrichment. 
In terms of the return mass fraction of stars $r$, the mass lost to remnants is $Z_\alpha\,\dot{M}_\star\,(1-r)$.  And, the outflows deplete mass at a rate $Z_\alpha \,\dot{M}_\star\,\eta$. (We assume the composition of outflows is the same as the interstellar medium.) With these definitions, Eq.~\ref{eq:mdot_1} is
\begin{equation}
    \dot{M}_\alpha= y_\alpha\,\dot{M}_\star - (1 + \eta - r)\,Z_\alpha\,\dot{M}_\star.
\end{equation}
Assuming an exponentially declining star formation history (\sfh{}) $\dot{M}_\star \propto e^{-t/\tau_{\rm sfh}}$, the equilibrium abundance is derived analytically by setting $\dot{Z}_\alpha=0$.
\begin{equation}\label{eq:z_eq}
    Z_\alpha^{\rm eq}(R) = \frac{y_\alpha}{1 + \eta(R) - r - \tau_\star / \tau_{\rm sfh}},
\end{equation}
where $\tau_{\star}$ is the star formation rate. ($\eta$ depends on $R$ to create a metallicity gradient. See also Section \ref{sec:vice} and Eq.~\ref{eq:eta_r}.)
In the special case of constant star formation, $\tau_{\rm sfh}\to\infty$, the denominator simplifies to $1+\eta-r$.

@@

\section{Results}
\subsection{Data Selection}

>>>>

% Abstract of the paper
\begin{abstract}
% context
C is an important element across astronomy; however, its origin remains poorly understood. 
% 
We aim to constrain the stellar yields of C through multi-zone Galactic chemical evolution models by comparing predictions with \apogee\ subgiants abundances.
% 
We find that \caafe\ is an emirical estimate of the delayed C sources, enabling us to estimate that \agb\ stars and \cc\ produce about 20\% and 80\% of C, respectively.  
The \caah\ trend instead represents the equilibrium abundances of C and Mg. 
We use the \caah\ trend to estimate the CCSNe C/Mg yield, determining that  $\Ycc/\Yoc = EQUATION$ when including \agb\ C. 
% misc
Our models are relatively independent of uniform scaling of yields and outflows, and alternate star formation histories. 
However, the stars which contribute to \agb\ C production and the \ia\ delay time distribution of Fe contribute uncertainties to our conclusions. 
% gas phase
While reliable gas-phase and low-metallicity measurments of C are challenging, we find that our model and a singlezone model with our recommended yields replicate the broad trends of \caah across different environments and metallicities.

@@

Carbon is a distinctive and well-studied element in astronomy. 
Formed in the cores of stars during He fusion, C is the lightest directly synthesized element after He, one of the only light elements formed in low-mass stars, and one the most abundant metals \citep[e.g.][]{jennifer19, KL14}.\footnotemark{}
C structurally changes the environments it pollutes---regulating stellar evolution, facilitating the formation of stars and planets, and forming the basis of earthly life.
% C/N as age indicator?
As such, understanding the origin of C has wide ranging implications. 
While we know both lower-intermediate-mass and high-mass stars produce C, the relative importance of each process is still unknown.

\footnotetext{By metallicity, we mean the (mass) fraction of any element which is not H or He, denoted by $Z$. For the sun, we take $Z_\odot=0.014$. }

Galactic chemical evolution (\gce) is a powerful tool, capable of uncovering the origins of the elements. 
Each enrichment process has characteristic chemical signatures and timescales, enabling us to reconstruct chemical histories.
Many previous works have used GCE in attemt to understand C abundances, whether through using theoretical stellar models \citep{DTS78, prantzos+18, chiappini+03} 
or understanding  observation \citep{tinsley79, HEK00, BF06, rybizki+17, berg+19, KKL20}.
(See also review in \citealt{romano22}.)
%
Every study agrees that C is produced by a combination of high-mass and \lims{}, different studies disagree on which process is dominant. 
For example CITES conclude \hms{} dominate wherease CITES conclude \lims{} contribute the majority of C.
C is also generally understood to have strong metallicity dependent CCSNe enrichment.

One of the primary uncertainties of GCE models are nucleosynthetic yields. Yield predictions -- the amounts of each chemical element stars produce --
are shaped by poorly understood processes, including mass loss, nuclear reaction rates, rotational mixing, convection, and explodability \citep{romano+10,KL14,ventura+13, LC18, emily+21}.
To better understand where C comes from and how it evolves, our aim is to combine \apogee\ observations and multi-zone models to develop observationally-consistant yields.
\cite{james+23} examined similar \gce{} models of N (which is closely related to C), finding that trends in N and O are explained by the metallicity dependence of N/O yields. \citet{james+23} determine that \agb\ N abundances roughly dependence on metallicity (i.e. $y_{\rm N}/y_{\rm O} \propto Z$). 
Here, we extend their models to C, deriving similar constraints on C/Mg yields. We assess which yield prescriptions reproduce Galactic abundance trends while investigating the impact of \gce{} model assumptions, such as the star formation history (\sfh{}) and outflow mass loading.

% \footnotetext{e.g. \citet{HEK00,PVT10,berg+12, berg+20, skillman+20, izotov+12, james2+15, dopita+16}.} 
We know, from observations of MW stars and extragalactic gas measurments, that C/O traces a banana shape in O/H (see Fig.~\ref{fig:gas_phase} and section~\ref{sec:gas}). 
C (and O) have been observed in local and extragalactic HII regions (CITES), dwarf galaxies \citep[e.g.]{berg+19}, halo and thick disc Milky Way stars \cite[e.g.][]{fabbian+09, nissen+14, lambert81, laird85, lambert86}, and damped Lyman-alpha systems. 
\footnote{DLAs; high redshift clouds of gas observed as absorption lines in yet higher redshift quasars}
The many observations of C and O paint a broadly consistant picture of a global trend.
At the very lowest metallicities, where $[{\rm O/H}]\lesssim-2$,\footnotemark{} observations suggest that C/O declines with metallicity.
From $-2 \lesssim {\rm [O/H]}\lesssim -1$, C/O is roughtly constant with metallicities. 
And, at higher metallicities $-1 \lesssim {\rm [O/H]}$, C/O increases with increasing metallicity.

@@

However, observations of C are not without challenges. 
As stars ascend the red giant branch (RGB) at the end of H core burning, C abundances are modified through \textit{first dredge up}\footnotemark{}. Because the effects of first dredge up are mass-dependent, C and N abundances are frequently used as age indicators for RGB stars \citep{MG15, martig16, hasselquist19, vincenzo+21}. Evolved RGB stars thus require modelling-dependent corrections to recover birth C and N abundances \citep[e.g.][]{vincenzo+21}. 
C is also challenging to determine in optical spectroscopy. 
C lines are relatively faint, and surveys such as GALAH struggle against low detection rates, potentially biasing sample measurments. 
(e.g. gaia-eso AS WELL \citealt{franchini+20}).
In the gas phase, HII regions are our best window into C abundances. Unfortunantly, C lacks strong collisional excitation lines, and recombination lines fall in the ultraviolet without nearby reference H lines \citep{skillman+20}. Recombination-line and collisional-excitation-line measurements furthermore disagree by a factor of \about{2} \citep{GR07}.

\footnotetext{First dredge up is when material from CNO processed core is mixed with the (convective) stellar envelop during the ascent onto the RGB. As a result, C is decreased and N is increased depending on the strength of this process and the envelop mass.}

    As our primary observational constraint, we use a sample of subgiant stars from the \apogee{} \citep{apogee17} selected by the criteria in Roberts et al. (2023, in prep; see Appendix \ref{sec:jack}).
According to stellar evolution theory and observations \citep{gilroy89, korn+07, lind+08, souto+18, souto19}, these stars have not yet experienced first dredge up yet have well-mixed envelopes. So, subgiant surface abundances most accurately represent their birth composition. 
INTRO TO \apogee\ takes high-resolution near-infrared spectroscopic measurments of stars, allowing accurate determination of chemical abundances and stellar parameters. In particular, near-infrared C abundances are more reliable than in optical spectroscopy (CITE??).
Fig.~\ref{fig:subgiants} shows the subgiant sample plotted in \caah\ and \caafe{}. [C/Mg] increases with metallicity, and [C/Mg] decreases with [Mg/Fe] at fixed [Mg/H]. 
Using the subgiant abundance trends, we will develop a model of the enrichment sources and evolution of C.

@@

Yields -- the amounts of each chemical element stars synthesise -- are central to studies of galactic chemical evolution (GCE). 
In this section, we compare literature yield predictions and discuss our fiducial yield choices. 
We focus on three nucleosynthetic pathways: asymptotic giant branch (\agb{}) stars, core collapse supernovae (\cc{}), and type Ia supernovae (\ia{}).
C is produced in both \agb{} and \cc{} stars.
We also use Mg and Fe as tracers of \cc{} and \ia{} enrichment respectively. 
O and Mg are produced almost entirely from \cc\ with metallicity-independent yields. In contrast, Fe is produced in similar amounts by \cc\ and \ia.

@@

To quantify yields, we define the stellar yield to be the fraction of a stars initial mass which is newly synthesized and released as a given element. 
For an element $X$ and star with mass $M$, the net-fractional stellar yield $\y$ is 
\begin{equation}
\y_{X} =  \Delta Z_X \frac{M_{\rm ejected}}{M_{\rm birth}}
\end{equation}
where $M_{\rm ejected}$ and $M_{\rm birth}$  are the total ejected mass and the birth mass of the star, and $\Delta Z_X$ is the change in $Z_X$ from the birth material to the ejected material of the star.%
\footnote{$Z_{X}$ represents the mass fraction of element $X$.}
%
For example, a 1\,\Mo\ star with $\y_{\C} = 10^{-3}$ will add $10^{-3}\,\Mo$ of new C to the interstellar medium. 
Also, note that yields may be negative if the material returned to the interstellar medium has a lower abundance $Z_X$ than the material the star was formed from.
Although per-star yields are necessary to compute \agb{} star enrichment rates in \gce{}  models, initial-mass-function (\imf) averaged yields are useful in interpreting their predictions. An IMF-averaged yield adds together the yields of stars of each mass, weighted by the fraction of stars of each mass (the \imf). 
Given the fractional yield $\y(M, Z)$ as a function of initial stellar mass $M$ and metallicity $Z$, the \imf-averaged yield is given by 
\begin{equation} \label{eq:imf-yield}
    y_{\rm X}(Z,t) = 
    \frac{
    \int_{M_{\rm min}(t)}^{M_{\rm max}} 
    \y_{\rm X}(M, Z)
    \frac{dN}{dM}\ M\ dM
}
{
    \int \frac{dN}{dM}\ M\ dM
}
\end{equation}
where ${dN}/{dM}$ is the \imf, $M_{\rm max}=100\,\Mo$ is the maximum stellar mass, and $M_{\rm min}(t)$ is the mass of stars with lifetime $t$.\footnotemark{}
To calculate the \imf-averaged net yields, we use the Versatile Integrator for Chemical Evolution code (\VICE).\footnotemark{}

We adapt the yield choices of elements besides C from \citet{james+21, james+23}.
Table \ref{tab:fiducial_mod} contains our fiducial yields. 
Following \citet{james+21, james+23}, we also take the \ia{} delay time distribution to be a $t^{-1.1}$ power-law with a minimum delay time of 140\,Myr, as suggested by the observations of \citet{maoz+12}.

\addtocounter{footnote}{-2}
    \stepcounter{footnote}\footnotetext{In our model, the mass-lifetime relation is
$\log \tau_M = 1.02 - 3.57\log M + 0.90 \left(\log M\right)^2$,
where $\tau_M$ is in Gyr, from \citealt{larson74}.
We use $t_{\rm end}=10\,$Gyr for total yields when $t$ is not used.}

@@

An \agb\ star is a low-mass ($\lesssim 8 M_{\sun}$) star during its final phase of evolution. Inside an \agb\ star, an inert CO core is surrounded by thermally-unstable He and H-burning shells below a convective envelope. 
In an \agb\ star, two competing processes determine the outcome of C production: \textit{third dredge up} and \textit{hot bottom burning}.  
Third dredge up accompanies thermal pulses in \agb\ stars, where material from the CO core is mixed with the envelope, increasing surface C abundances \citep{KL14}. The C yields of the star are increased as this C-enhanced envelope is released to the interstellar medium. 
Hot bottom burning is the activation of the CNO cycle\footnotemark{}
at the bottom of the convective envelope when $T\gtrsim 50\,{\rm MK}$. Because the $^{14}$N proton capture is the slowest component of the CNO cycle, the CNO cycle converts nearly all \C[12] into $^{14}$N \citep{solar-fusion}.

@@

Hot bottom burning and third dredge-up result in mass-dependent C yields. 
Stars less than \about{2}\,\Mo do not experience third dredge-up. As a result, these stars C abundances are only affected by first dredge-up, resulting in little change to C yields or slight destruction of C.
Between \about{2} and \about{5} \Mo, third dredge up becomes important, enriching the outer layers with C, making these stars the most important producers of C. 
In \agb\ stars more massive than \about{5}\,\Mo, both hot bottom burning and third dredge up occur; however, hot bottom burning is much more efficient, resulting in significant \C[12] destruction.

@@

Fig.~\ref{fig:y_agb} compares the stellar \agb\ C yields for these four models.
Most models agree on the qualitative shape of the net fractional \agb\ C yield.
Stellar yields peak between masses of about 2--4 \Mo and decline as stars become more or less massive. As metallicity increases, the total \agb\ C yield decreases. The mass of peak C yields also increases slightly with metallicity. Metal poor stars dredge up more material due to the decreased power of the CNO cycle, resulting in increasing carbon yields with decreasing metalicity \citep{ventura+13}.

Fig.~\ref{fig:agb-ssp}, on the left, shows the total production of C by \agb\ stars in a \ssp{} at an age $t$, i.e. $\y_{\rm C}(Z_\odot, t)$. 
As the mass range $2\,\Mo\lesssim M \lesssim 4\,\Mo$ is most important for C production, about half of C production occurs before \about{1}\,Gyr, similar to \ia\ Fe. 
\kx{} and \kxvi{} weight C production more heavily towards high-mass \agb\ stars resulting in a faster enrichment delay time, whereas the \cxi\ and \vxiii\ models predict a slightly longer timescale of \about{1}\,Gyr. In any case, little to no C is produced more than 2\,Gyr after a star formation event. Fe production, in contrast, continues steadily for 10\,Gyr.

The right panel of Fig.~\ref{fig:agb-ssp} shows \imf-averaged C yields for each \agb\ model as a function of metallicity.
\vxiii{} differs in that it shows a non-monotonic metallicity dependence. However, this effect is only for models with $\log Z/Z_\odot \lesssim -1$.
Otherwise, models differ only in their yield normalization and metallicity dependence. 
All yield models span a range of \about{2} for a given metallicity.
For example, all four models predict $y_\text{C}^\text{agb}(\Zo)$ to be between 0.004 and 0.008 at solar metallicity. The models also differ slightly in the strength of metallicity dependence, with about a factor of 3 discrepency (see Table~\ref{tab:alpha_agb}).

@@

Massive stars form $^{12}$C in their cores through the triple--$\alpha$ process.\footnote{The 3-$\alpha$ process fuses 3 He nuclei into O.} 
However, only C ejected through supernovae and stellar winds contributes to the yield. 
While there are many stellar models providing predictions of \cc{} yields, the results of these models are highly uncertain due to the complexity of stellar modeling.

@@

Both rotation and explodability introduce substantial variations. The \cite{LC18} models include rotation, showing that variations in the rotational velocity of the star can dramatically increase the magnitude and metallicity dependence of $\Ycc$. Rotation induces more mixing allowing the CO core to grow larger and contributes to wind losses. As we will later show, \cc\ C production needs to be strongly metallicity-dependent at $Z/Z_\odot \approx 1$, which is consistent with the \cite{LC18} rapidly rotating models.
Assumptions about the explodability landscape affect C and Mg production. As fewer high-mass stars explode, both C and Mg yields decrease, but Mg yields decrease more as more C is lost to winds, so [C/Mg] increases with decreasing explodability.

O and Mg are both $\alpha$-elements, light elements produced in \cc{} through He-nuclei fusion. 
As we focus on constraining relative yields, we neglect O and Mg yield variations in the main text (excluding the uniform scaling of yields and massloading in Section~\ref{sec:outflows}). There is substantial variation in predicted Mg yields. 
Most models predict relatively flat trends metallicity (even with rotation as in \citealt{LC18}). 
However, the variation is significant and our adopted $\Yoc$ yield is much higher than most models, but O and Mg yields of \cc\ models do not fully match observations. \cc\ models underpredict $[{\rm Mg}/{\rm O}]$, and the reason why is unknown (see e.g. \citealt{emily+21}). Here, we assume ${\rm [O/Mg]} = 0$, which nevertheless consistent with \apogee{} observations \citep{weinberg+19, weinberg+22}.

@@

Here, we assume a simple \textit{one-zone} chemical evolution model \cite[e.g.][]{tinsley80, pagel09, matteucci21}.  Newly produced metals are homogeneously and instantaneously mixed, so this model evolves as a single, spatially-uniform chemical zone.
We define $M_{X}$ to be the gas mass of element $X$, $\dot{M}_\star$ to be the star formation rate (in $\Mo\,{\rm yr}^{-1}$), and $\eta$ to be the mass loading factor $\eta\equiv\dot{M}_{\rm outflow}/\dot{M}_\star$ (representing the strength of outflows). A \ssp\ returns a fraction $r$ of their birth mass to the interstellar medium, due to ejected stellar envelopes.%
\footnote{$r\approx0.4$ for a \citealt{kroupa01} \imf.}
Here, we will use $\alpha$ as a representative $\alpha$-element, such as O or Mg. 
Given the \imf-averaged yield $y_\alpha$, the rate of change in the gas-phase mass of $\alpha$ is a sum of sources and sinks,
\begin{equation} \label{eq:mdot_1}
    \dot{M}_\alpha =  y_\alpha\,\dot{M}_\star - \dot{M}_{\alpha,\,\text{remnants}} - \dot{M}_{\alpha,\,\text{outflows}}
\end{equation}
where $y_\alpha\,\dot{M}_{\star}$ describes \cc\ enrichment. 
In terms of the return mass fraction of stars $r$, the mass lost to remnants is $Z_\alpha\,\dot{M}_\star\,(1-r)$.  And, the outflows deplete mass at a rate $Z_\alpha \,\dot{M}_\star\,\eta$. (We assume the composition of outflows is the same as the interstellar medium.) With these definitions, Eq.~\ref{eq:mdot_1} becomes
\begin{equation}
    \dot{M}_\alpha= y_\alpha\,\dot{M}_\star - (1 + \eta - r)\,Z_\alpha\,\dot{M}_\star.
\end{equation}
Assuming an exponentially declining star formation history (\sfh{}) $\dot{M}_\star \propto e^{-t/\tau_{\rm sfh}}$, the equilibrium abundance is derived analytically by setting $\dot{Z}_\alpha=0$.
\begin{equation}\label{eq:z_eq}
    Z_\alpha^{\rm eq}(R) = \frac{y_\alpha}{1 + \eta(R) - r - \tau_\star / \tau_{\rm sfh}},
\end{equation}
where $\tau_{\star}$ is the star formation rate. ($\eta$ depends on $R$ to create a metallicity gradient. See also Section \ref{sec:vice} and Eq.~\ref{eq:eta_r}.)
In the special case of constant star formation, $\tau_{\rm sfh}\to\infty$ and the denominator simplifies to $1+\eta-r$.

@@

\section{Results}
